Recovering lost 21-cm radial modes for cross correlation with the CMB and photo-$z$ galaxies via cosmic tidal reconstruction

### 2018-02-21 ms.tex
Sep 22th
Feb 21th

\documentclass[aps,prl,twocolumn,showpacs,superscriptaddress,groupedaddress,nofootinbib,floatfix]{revtex4}  % for review and submission
%\documentclass[aps,preprint,showpacs,superscriptaddress,groupedaddress]{revtex4}  % for double-spaced preprint
\usepackage{graphicx}  % needed for figures
\usepackage{dcolumn}   % needed for some tables
\usepackage{bm}        % for math
\usepackage{amsmath,amssymb}   % for math
\usepackage{aas_macros}
\usepackage{multirow}
\usepackage{color}
\usepackage{verbatim}
\usepackage{times}
\usepackage{url}
\usepackage{hyperref}

% avoids incorrect hyphenation, added Nov/08 by SSR
\hyphenation{ALPGEN}
\hyphenation{EVTGEN}
\hyphenation{PYTHIA}

\newcommand{\mr}{\mathrm}
\newcommand{\tcb}{\textcolor{blue}}

\begin{document}
% The following information is for internal review, please remove them for submission
\widetext
% the following line is for submission, including submission to the arXiv!!
%\hspace{5.2in} \mbox{Fermilab-Pub-04/xxx-E}

\title{Recovering Lost 21cm Radial Modes via Cosmic Tidal Reconstruction}

\author{Hong-Ming Zhu} 
\affiliation{Key Laboratory for Computational Astrophysics,  National Astronomical Observatories, Chinese Academy of Sciences, 20A Datun Road, Beijing 100012, China}
\affiliation{University of Chinese Academy of Sciences, Beijing 100049, China}

@@

\author{Yu Yu}
\affiliation{Key Laboratory for Research in Galaxies and Cosmology,
Shanghai Astronomical Observatory, Chinese Academy of Sciences,
80 Nandan Road, Shanghai 200030, China}

@@

\begin{abstract}
21 cm intensity mapping has emerged as a promising technique to map the 
large-scale structure of the Universe, at redshifts $z$ from 1 to 10.
Unfortunately, many of the key cross correlations with photo-$z$ galaxies
and the CMB have been thought to be impossible due to foreground
contamination for radial modes with small wavenumbers. 
These modes are usually subtracted in the foreground subtraction process.
We recover the lost 21 cm radial modes via cosmic tidal reconstruction and 
find more than 60\% cross-correlation signal at $\ell\lesssim100$ and 
even more on larger scales can be recovered from null. 
The tidal reconstruction method opens up a new set of possibilities to 
probe our Universe and is extremely valuable not only for 21 cm surveys
but also for CMB and photometric redshift observations.
\end{abstract}

\pacs{}
\maketitle

%\section{\label{sec:level1}First-level heading}
% sections are not used for PRL papers

{\it Introduction.}---The current and future cosmological observations aim
to map a large fraction of the Universe with unprecedent precision, varying
from LSST \cite{2009:lsst}, Euclid \cite{2012:euclid}, 
Planck \cite{2015arXiv150201582P}, CMB-S4 \cite{2014ApJ...788..138W}, and etc. 
In addition to these ones, 21 cm intensity mapping has emerged as
a powerful probe of the 
cosmological large-scale structure \cite{2008:21cm,chang}.
However, the astrophysical foregrounds from galactic and extra-galactic
synchrotron and free-free emissions are stronger than the cosmological 21 cm 
signals by many orders of magnitude. These foregrounds are expected to be 
spectrally smooth, which means they would contaminate small radial
modes, i.e., all modes with low $k_\parallel$.
While there are not many modes with small $k_\parallel$, many other cosmic
observations with broad window functions along the radial direction only 
probe these modes, such as weak lensing, photo-$z$ galaxies, 
integrated Sachs-Wolf effect, kinetic Sunyaev-Zel'dovich effect, and etc.  
Thus, 
%this makes it very hard to cross correlate these cosmic probes with 
%21 cm intensity mapping surveys 
the cross correlations of 21 cm surveys with these cosmic probes are dominated 
by foreground contamination \cite{2007ApJ...660.1030F,2008MNRAS.384..291A}. 
To solve this problem is of great importance for both 21 cm 
intensity mapping surveys and other key cosmic surveys.

Recently a new method called {\it cosmic tidal reconstruction} has been 
developed \cite{2012:pen,2015:zhu}. It can reconstruct the large-scale 
tidal field and hence large-scale density field from the alignment of 
small-scale cosmic structures.
The modes with small $k_\parallel$ and large $k_\perp$ are well 
reconstructed \cite{2015:zhu}, which are exactly those lost in the foreground 
subtraction in 21 cm experiments. 
This technique enables the reconstruction of small radial modes, 
which provides important radial information essential for cross correlating 
with CMB and other cosmic probes.

%which are essential for CMB and other cross correlations.

In this Letter we study how to use cosmic tidal reconstruction to 
reconstruct the lost large-scale radial modes, and further cross correlate with
CMB lensing, photo-$z$ galaxies and ISW effect. 
We find such reconstruction technique recovers more than 60\% cross-correlation 
signal at $\ell\lesssim100$ from nothing and even better on larger scales. 
This provides a new way to probe the origin and evolution of the Universe.

{\it Cosmic tidal reconstruction.}---Cosmic tides is a new way to view the
tidal effect of gravity on the structure of matter clustering \cite{2012:pen}.
The large-scale density field can be reconstructed accurately from the 
anisotropic tidal distortions of the locally-measured matter power spectrum
\cite{2012:pen,2015:zhu}.
The basic idea of purely transverse tidal reconstruction has been proposed in
Ref. \cite{2012:pen} and further expanded in Ref. \cite{2015:zhu}.
Here, we briefly discuss the idea and summarize the process of cosmic
tidal reconstruction.

The evolution of small-scale density perturbations is modulated by 
long-wavelength perturbations \cite{2014:tidal}. 
The anisotropic distortions in the local small-scale matter power specturm,
$\propto\hat{k}^i\hat{k}^jt_{ij}^{(0)}$, arise from the coupling of small-scale
density fluctuations with the large-scale tidal field $t_{ij}$, where 
$t_{ij}=\Phi_{L,ij}-\delta_{ij}\nabla^2\Phi_L/3$, $\hat{\ }$ denotes the 
unit vector and superscript (0) denotes some ``initial'' time. 
The traceless tidal field $t_{ij}$ can be decomposed into 5 independently 
observable components \cite{2015:zhu}.
The two transverse shear terms, 
$\gamma_1=(\Phi_{L,11}-\Phi_{L,22})/2$ and $\gamma_2=\Phi_{L,12}$,
which describe quadrupolar distortions in the tangential plane perpendicular
to the line of sight, are less affected by peculiar velocities and used to 
perform reconstruction \cite{2015:zhu}.
%Since 
%Here, $\gamma_1=(\Phi_{L,11}-\Phi_{L,22})/2$ and $\gamma_2=\Phi_{L,12}$
%only involves derivatives in the tangential plane, the changes in $\gamma_1$
%and $\gamma_2$ due to the redshift space distortions are expected to be second
%order effects. 
The two gravitational tidal shear fields $\gamma_1$ and $\gamma_2$
can be converted to the two-dimensional (2D) convergence field, 
$\kappa_\mr{2D}=(\Phi_{L,11}+\Phi_{L,22})/2$, using
\begin{eqnarray}
\label{eq:kappa2d}
\kappa_{\mr{2D},11}+\kappa_{\mr{2D},22}=
(\gamma_{1,11}-\gamma_{1,22}+2\gamma_{2,12}).
\end{eqnarray}
The three-dimensional (3D) convergence $\kappa_\mr{3D}=\nabla^2\Phi_L/3\propto\delta_L$, which gives
the large-scale density field, is given by 
\begin{eqnarray}
\label{eq:kappa3d}
\kappa_{\mr{3D},11}+\kappa_{\mr{3D},22}=\frac{2}{3}\nabla^2\kappa_\mr{2D}.
\end{eqnarray}
Since only two transverse tidal shear fields $\gamma_1(\bm{x})$ and 
$\gamma_2(\bm{x})$ are used, the changes of the large-scale density field along
the line of sight are inferred from the variations of $\gamma_1$ and $\gamma_2$
along the $x_\parallel$ axis. The error of $\kappa_\mr{3D}$ is 
\begin{eqnarray}
\sigma_{\kappa_\mr{3D}}(\bm{k})\propto\big({k^2}/{k_\perp^2}\big)^2,
\end{eqnarray}
which is anisotropic in $k_\perp$ and $k_\parallel$ \cite{2015:zhu}.
The reconstruction works best for modes with low $k_\parallel$ and high 
$k_\perp$, which can not be obtained from 21 cm surveys and contribute 
substantially to cosmological observables from other surveys mentioned above.
Thus, cosmic tidal reconstruction provides a new way to recover the lost 
radial modes and to improve the cross correlations.

The tidal reconstrction works as follows. The first step is to convolve
the density field with a Gaussian kernel, $S(\bm{k})=e^{-k^2R^2/2}$, which
filters out the small-scale nonlinear structures.
Here, we still take $R=1.25\ \mr{Mpc}/h$ \cite{2012:pen,2015:zhu}. The
next step is to gaussianize the smoothed density field by taking a logarithmic
transform or mapping the density fluctuations into a Gaussian distribution. 
In the following reconstructions, we adopt the latter method since after the
simulated
foreground subtraction, some of the density contrasts become smaller than $-1$,
which makes it hard to take the logarithmic transform, $\mathrm{ln}(1+\delta)$.
The gravitational tidal shear fields can be estimated by applying quadratic 
tidal shear estimators $\hat{\gamma}_1$ and $\hat{\gamma}_2$ to the density 
field as in 21 cm lensing reconstruction \cite{2008:lu}. 
Then the 3D tidal convergence field $\kappa_\mr{3D}$ is given by the linear 
combination of tidal shear fields using Eq. (\ref{eq:kappa2d}) and 
Eq. (\ref{eq:kappa3d}).
The reconstructed noisy field $\kappa_\mr{3D}$ is related to the original 
density field as 
\begin{eqnarray}
\kappa_\mr{3D}(k_\perp,k_\parallel)=b(k_\perp,k_\parallel)
\delta(k_\perp,k_\parallel)+n(k_\perp,k_\parallel),
\end{eqnarray}
where $b=P_{\kappa_\mr{3D}\delta}/P_{\delta}$ is the bias factor and $n$ is 
the noise of reconstruction \cite{2015:zhu}. 
The reconstructed clean field is given by 
\begin{eqnarray}
\label{eq:kapc}
\hat{\kappa}_c=({\kappa_\mr{3D}}/{b})W,
\end{eqnarray}
where the Wiener filter $W(k_\perp,k_\parallel)=
P_\delta/(P_{\kappa_\mr{3d}}/b^2)$.

{\it Simulation setup.}---We further explore this idea with numerical 
simulations. We employ an ensemble of six $N$-body simulations from the
$\mr{CUBEP}^3\mr{M}$ code \cite{2013:code}. 
Each simulation includes $1024^3$ particles in a $(1.2\mr{Gpc}/h)^3$ box. 
%with following cosmological parameters: Hubble parameter $h=0.678$, baryon
%density $\Omega_{b}=0.049$, dark matter density $\Omega_{c}=0.259$,
%amplitude of primordial curvature power spectrum $A_s=2.139\times10^{-9}$ at 
%$k_0=0.05\;\mr{Mpc}^{-1}$ and scalar spectral index $n_s=0.968$.
In the following analysis we use outputs at $z=1$.

\begin{figure}[tbp]
\begin{center}
\includegraphics[width=0.48\textwidth]{1.000cc2d.eps}
\end{center}
\vspace{-0.7cm}
\caption{(Top) The cross-correlation coefficient of the  
foreground subtracted density field with the original density field.
%$r_{\delta_{fs}\delta}=P_{\delta_{fs}\delta}/\sqrt{P_{\delta_{fs}}P_\delta}$. 
(Bottom) The cross-correlation coefficient of the reconstructed field with 
the original density field. 
%$r_{\kappa_{c}\delta}=P_{\kappa_{c}\delta}/\sqrt{P_{\kappa_{c}}P_\delta}$. 
The lost radial modes appear again after reconstruction.
The results are for $R_\parallel=60\ \mr{Mpc}/h$ and $k_c=0.5\ h/\mr{Mpc}$.}
\label{fig:ratio}
\end{figure}

We could approximately use dark matter to represent 21 cm source distributions.
This is a good approximation since the neutral hydrogen traces the total mass
distribution fairly well at low redshifts. We simply assume the experimental
noise to be zero above a cut off scale and infinity below the cut off scale.
This is a reasonable approximation for a filled aperture experiment, which
has good brightness sensitivity and an exponetially growing noise at small 
scales.
We choose this scale to be $k_c=0.5\ h/\mr{Mpc}$, which corresponds to 
$\ell=1150$ at $z=1$. This is realistic for the ongoing 21 cm experiments like
CHIME \cite{2014SPIE.9145E..22B}, Tianlai \cite{2015ApJ...798...40X}, 
HIRAX \cite{HIRAX}, and etc.

We are not going to provide a detailed 21 cm data reduction process in this 
Letter, so we simply  use a high-pass filter 
$W_{fs}(k_\parallel)=1-e^{-k_\parallel^2R_\parallel^2/2}$ to simulate the 
foreground subtraction. We show results for the two different scales 
$R_\parallel=60\ \mr{Mpc}/h$ and $R_\parallel=15\ \mr{Mpc}/h$, which give
$W_{fs}=0.5$ at $k_\parallel=0.02\ \mr{Mpc}/h$ and 
$k_\parallel=0.08\ \mr{Mpc}/h$, respectively. The former is an optimal case, 
i.e., we remove modes for $k_\parallel\lesssim0.02\ \mr{Mpc}/h$ while the latter
is already achieved in the current 21 cm observations 
\cite{2013ApJ...763L..20M,2013MNRAS.434L..46S}.

The observed 21 cm field after foreground subtraction is given by 
\begin{eqnarray}
\delta_{fs}(\bm{k})=\delta(\bm{k})W_{fs}(k_\parallel)\Theta(k_c-k),
\end{eqnarray}
where $\delta(\bm{k})$ is the density field from simulations, $W_{fs}$ accounts
the effect of foreground subtraction and $\Theta(x)$ is the step function 
which equals $1$ for $x\ge0$ and $0$ otherwise.
Then we get the reconstructed clean field $\kappa_c$ from $\delta_{fs}$ via
cosmic tidal reconstruction. In Fig. \ref{fig:ratio}, the upper panel shows the 
cross-correlation coefficient of $\delta_{fs}$ with $\delta$ and the lower 
panel shows the cross-correlation coefficient of $\kappa_c$ with $\delta$ for 
$R_\parallel=60\ \mr{Mpc}/h$. 
We find the lost radial modes appear again in the reconstructed field.

{\it Cross-correlation signals.}---To show how much the cross-correlation signal
is recovered by cosmic tidal reconstruction and estimate the detectability of 
the cross correlation, we need to generate the lensing convergence field,
the angular distribution of photo-$z$ galaxies and the temperature fluctuation
due to the ISW effect using $N$-body simulations.

\begin{figure}[tbp]
\begin{center}
\includegraphics[width=0.48\textwidth]{fig1b.eps}
\end{center}
\vspace{-0.7cm}
\caption{The cross-correlation coefficients of 21 cm with different 
observations for both $R_\parallel=60\ \mr{Mpc}/h$ and $15\ \mr{Mpc}/h$ with 
$k_c=0.5\ h/\mr{Mpc}$. Cosmic tidal reconstruction recovers about 60\% signal 
at $\ell\sim100$ and much better results are obtained on larger scales. 
The error bars are estimated using the bootstrap resampling method.}
\label{fig:cc}
\end{figure}

%\begin{figure}[tbp]
%\begin{center}
%\includegraphics[width=0.48\textwidth]{f8.eps}
%\end{center}
%\vspace{-0.7cm}
%\caption{The cross correlation signals between 21cm and other cosmic probes.}
%\end{figure}

(1) CMB lensing.---The weak lensing convergence is a weighted projection
of the dark matter density fluctuations along the line of sight to the last 
scattering surface,
\begin{eqnarray}
\kappa(\bm{\theta})=\int_0^{\chi_s}d\chi 
W_\kappa(\chi)\delta(\chi\bm{\theta},\chi)\ ,
\end{eqnarray}
where the lensing kernel
\begin{eqnarray}
W_\kappa(\chi)=\frac{3\Omega_{m0}H_0^2\chi(\chi_s-\chi)}{2a(\chi)\chi_s}\ ,
\end{eqnarray}
with $\chi_s=\chi(z_s=1090)$.
To generate the convergence field from simulation data, we first squeeze the 
simulation box to a 2D plane, then multiply the $W(\chi(z=1))$, as the 
lensing weight is a slowly varying function for the case of CMB lensing.
Then we obtain the lensing convergence field contributed by the simulated 
density field.

(2) Photo-$z$ galaxies.---We calculate the projected galaxy density field at 
$z\sim 1$ with usual photo-$z$ bin width of $0.2$, i.e., $z_\mr{p}\in(0.9,1.1)$.
We adopt the galaxy distribution characterized by 
$n(z)\propto z^{\alpha}\mr{exp}[-(z/z^{*})^\beta]$
with $\alpha=2$, $z^*=0.5$, $\beta=1$
and assume the photo-$z$ scatter $\mathcal{P}(z_\mr{p}|z)$ is perfectly known to be in a Gaussian form with photo-$z$ error $\sigma_z=0.05(1+z)$.
The 2D angular galaxy distribution is given by
\begin{eqnarray}
\delta_\mr{2D}(\bm{\theta})=\int_0^\infty dzW_\mr{p}(z)
\delta(\chi(z)\bm{\theta},\chi(z))\ ,
\end{eqnarray}
where the window function 
\begin{eqnarray}
W_\mr{p}(z)\propto\int_{0.9}^{1.1} n(z) \mathcal{P}(z_\mr{p}|z) dz_\mr{p}\ 
\end{eqnarray}
with normalization $\int W_\mr{p}(z)dz=1$.

>>>>

\documentclass[prd,superscriptaddress,floatfix,notitlepage,nofootinbib,reprint]{revtex4-1}

\usepackage{graphicx}  
\usepackage{dcolumn}   
\usepackage{bm}        
\usepackage{amsmath,amssymb}
\usepackage{aas_macros}
\usepackage{multirow}
\usepackage{color}
\usepackage{verbatim}
\usepackage{url}
\usepackage{hyperref}

\hyphenation{ALPGEN}
\hyphenation{EVTGEN}
\hyphenation{PYTHIA}

\newcommand{\mr}{\mathrm}
\newcommand{\kpa}{k_\parallel}
\newcommand{\kpe}{k_\perp}
\newcommand{\bmk}{\bm{k}}
\newcommand{\bmx}{\bm{x}}
\newcommand{\bea}{\begin{equation}}
\newcommand{\eea}{\end{equation}}

\begin{document}
\widetext

\title{Recovering lost 21-cm radial modes for cross correlation with the CMB and
photo-$z$ galaxies via cosmic tidal reconstruction}

\author{Hong-Ming Zhu}
\email[]{hmzhu@berkeley.edu}
\affiliation{Key Laboratory for Computational Astrophysics, National Astronomical Observatories, Chinese Academy of Sciences, 20A Datun Road, Beijing 100012, China}
\affiliation{University of Chinese Academy of Sciences, Beijing 100049, China}
\affiliation{Berkeley Center for Cosmological Physics and Department of Physics,
University of California, Berkeley, California 94720, USA}

@@

\author{Yu Yu}
\affiliation{Department of Astronomy, Shanghai Jiao Tong University, 800 Dongchuan Road, Shanghai 200240, China}

@@

\begin{abstract}
21-cm intensity mapping has emerged as a promising technique to map the 
large-scale structure of the Universe, at redshifts $z$ from 1 to 10.
Unfortunately, many of the key cross correlations with the cosmic microwave 
background and photo-$z$ galaxies have been thought to be impossible due to 
foreground contamination for radial modes with small wave numbers. 
In this paper, we apply tidal reconstruction to the simulated 21-cm fields
and recover the lost large-scale radial modes successfully.
We estimate the detectability of the cross correlation signals and find them
can be detected at high significance.
The tidal field reconstruction method opens up a new set of possibilities to 
probe the Universe and is extremely valuable not only for 21-cm surveys
but also for cosmic microwave background and photometric redshift observations.
\end{abstract}

\maketitle

\section{Introduction}
The current and future cosmic surveys aim to map a large fraction of the 
Universe with unprecedented precision by observing the large-scale structure 
(e.g. SDSS \cite{2017SDSS}, DES \cite{2016DES}, PFS \cite{2014PFS}, DESI 
\cite{2016DESI}, LSST \cite{2009LSST}, Euclid \cite{2012Euclid}) and the cosmic
microwave background (CMB) (e.g. Planck \cite{2016Planck}, SPT-3G \cite{2014SPT-3G}, Advanced ACTPol \cite{2016AdvACTPol}, CMB-S4 \cite{2016CMB-S4}). 
Precision measurements of cosmological parameters from the autocorrelations
of large-scale structure and CMB observations and the cross correlations 
between different observations can improve constraints on the properties of 
dark energy, modifications to general relativity, neutrino masses, and 
primordial non-Gaussianities substantially.
In addition to these observation methods, 21-cm  intensity mapping has 
emerged as a powerful method to map the large-scale structure of the Universe 
\cite{2008IM,2008IM2,2010IM}. 
Instead of resolving millions of individual galaxies, the 21-cm intensity 
mapping technique measures the large-scale structure by detecting the aggregate 21-cm emission of neutral hydrogen from many galaxies in large voxels. 
The redshifted 21-cm emission line which provides the redshift information can 
be resolved exquisitely in the frequency domain.
Therefore, this allows radio telescopes to conduct rapid and efficient surveys
of large volumes of the Universe.
The ongoing and upcoming 21-cm surveys including CHIME \cite{2014CHIME}, HIRAX 
\cite{2016HIRAX}, Tianlai \cite{2015Tianlai}, BINGO \cite{2013BINGO}, FAST 
\cite{2016FAST}, MeerKAT \cite{2017MeerKAT}, and SKA \cite{2015SKA} can improve 
the baryon acoustic oscillations (BAO) measurements by observing a larger cosmic
volume at higher redshifts compared to current galaxy surveys.

The primary challenge for 21-cm intensity mapping experiments is the presence
of the astrophysical foregrounds from galactic and extra-galactic synchrotron 
emissions, which are three orders of magnitude brighter than the cosmological
21-cm signals. 
The synchrotron foregrounds are known to be spectrally smooth in the frequency
domain where the redshifted 21-cm signals from different redshifts fluctuate at 
different frequencies.
In principle, the foregrounds only impact the long wavelength density 
fluctuations along the line of sight, i.e., the modes with small $\kpa$ in 
Fourier space \cite{2003FG,2006FG}. 
However, the instrumental effects (e.g. spectral response, calibration, etc) 
further lead to an unsmooth foreground component, often referred as the 
foreground wedge at the low $\kpa$ and high $\kpe$ area in Fourier space 
\cite{2010Wedge,2012Wedge,2012Wedge2,2013Wedge,2015Wedge}.
The synchrotron foregrounds can be cleaned by exploiting their smooth spectral 
structure \cite{2011FG,2012FG,2014FG,2015FG}.
As demonstrated in Ref. \cite{2015FG}, the foregrounds can be cleaned well 
below the foreground wedge with the precise calibration of the instrument, 
leaving only $\kpa<0.02\ h\mr{Mpc}^{-1}$ modes unaccessible.

However, while there are not many Fourier modes at $\kpa<0.02\ h\mr{Mpc}^{-1}$, 
many other cosmological observations such as weak lensing, photometric-redshift
galaxies, and integrated Sachs-Wolf (ISW) effect can only probe these modes.
These observations involve a broad window function along the line of sight and
measure the projected modes, i.e., the modes with small $\kpa$, which are all
contaminated by the foreground emissions in 21-cm intensity mapping 
observations.
Therefore, the proposed cross correlation of 21-cm intensity mapping with weak
gravitational lensing \cite{2015xLensing&Photoz,2016xLensing,2016xLensing2}, 
photo-$z$ galaxies \cite{2015xLensing&Photoz,2015xPhotoz,2015xPhotoz2,2017xPhotoz,2017xPhotoz&ISW,2017xPhotoz2}, and ISW effect \cite{2017xPhotoz&ISW} would 
be severely degraded in the presence of foregrounds.
Recovering the lost large-scale radial modes for cross correlations is thus 
crucial in order to fully exploit the 21-cm intensity mapping experiments.
The cross-correlation measurements will benefit other observations as well,
since the cross correlations are expected to be more robust to the systematics
than autocorrelations from individual experiments. 
It also enables the use of sample variance cancellation technique \cite{2009MT}
to measure cosmological parameters \cite{2015xPhotoz,2015xPhotoz2,2017xPhotoz}.

Recently a new method called cosmic tidal reconstruction has been developed
\cite{2012Tides,2016Tides}. 
The small-scale density fluctuations are significantly affected by the 
large-scale density field as a consequence of nonlinear mode coupling.
The large-scale tidal shear field causes anisotropic distortions of the locally
measured small-scale matter power spectrum. 
Such local anisotropic tidal distortions can be exploited to reconstruct the 
large-scale tidal shear and hence density fields.
The reconstruction of gravitational tidal shear from local small-scale matter 
power spectrum is described by the same formulation as the reconstruction of 
gravitational lensing induced shear. 
As shown in Ref. \cite{2016Tides}, the density modes with small $\kpa$ 
and large $\kpe$ are well reconstructed, with cross-correlation coefficient
close to $1$ for reconstruction with the full dark matter density field.
These reconstructed density modes are exactly those lost in the foreground 
subtraction of 21-cm experiments. 
The tidal reconstruction technique enables the reconstruction of lost 21-cm 
radial modes, which provides important radial information essential for cross 
correlating with the CMB and photometric observations.

In this paper, we apply cosmic tidal reconstruction to the simulated 21-cm 
field with low $\kpa$ foreground modes subtracted. The small $\kpa$ radial 
modes are recovered successfully after tidal reconstruction.
Then we cross correlate the reconstructed 21-cm field with the simulated CMB 
lensing, photo-$z$ galaxy and ISW effect fields and estimate the detectability
of the cross-correlation signals with the current 21-cm experiments.
The tidal field reconstruction method provides us a new way to study the 
large-scale matter distribution in the Universe through cross correlations 
and has profound implications for the current and future 21-cm experiments.

The paper is organized as follows.
In Sec. \ref{sec:rec}, we introduce the cosmic tidal reconstruction.
In Sec. \ref{sec:sim}, we apply tidal field reconstruction to the simulated 
foreground subtracted 21-cm density field and shows the reconstruction results.
Section \ref{sec:ccs} shows the cross-correlation signal recovered after
reconstruction and estimates the detectability with current 21-cm experiments.
We discuss further improvements and future applications in Sec. \ref{sec:dis} 
and conclude in Sec. \ref{sec:cls}.

\section{Cosmic tidal reconstruction}
\label{sec:rec}
The large-scale density field can be reconstructed accurately from the 
anisotropic tidal distortions of the locally-measured matter power spectrum
\cite{2012Tides,2016Tides}.
The basic idea of purely transverse tidal reconstruction has been proposed in 
Ref. \cite{2012Tides} and further expanded in Ref. \cite{2016Tides}.
In this section, we briefly discuss the physical idea and outline the 
operational procedure of the tidal field reconstruction. More details of this 
reconstruction method are presented in Ref. \cite{2016Tides}.

\subsection{Cosmic tides}

The evolution of small-scale density perturbations is modulated by long 
wavelength perturbations during nonlinear structure formation. 
The gravitational coupling of a long wavelength tidal field with small-scale
density fluctuations has been studied extensively \cite{2014Tides}.
The leading-order observable of a long wavelength density perturbation on 
small-scale density perturbations is described by the large-scale tidal field,
\bea    
t_{ij}=\Phi_{L,ij}-\delta_{ij}\Phi_{L,kk}/3,
\eea
where $\Phi_L$ is the long wavelength gravitational potential sourced by the 
long wavelength density perturbation $\delta_L$. Here $\Phi_{L,ij}$ denotes
partial derivatives of $\Phi_L$ to $x^i$ and $x^j$.
Note that we have projected out the trace of $\Phi_{L,ij}$, which corresponds 
to the local mean density.
Since the change of shape is more robust than the change of number density, we 
shall focus on the gravitational tidal shear, i.e., the traceless tidal field.
The locally observed matter power spectrum in the presence of the large-scale
tidal field $t_{ij}$ can be calculated using Lagrangian perturbation theory 
and is given by
\bea
\label{eq:pk}
P(\bmk,\tau)|_{t_{ij}}=P(k,\tau)+\hat{k}^i\hat{k}^jt^{(0)}_{ij}P(k,\tau)f(k,\tau),
\eea
where $\tau$ is the conformal time, $P(k,\tau)$ is the isotropic linear power 
spectrum, $\hat{\bmk}$ is the unit vector, the superscript $(0)$ denotes the 
initial time defined in perturbation calculation. 
The coupling of the large-scale tidal field to small-scale density fluctuations
is described by the tidal coupling coefficient
\bea
f(k,\tau)=2\alpha(\tau)-\beta(\tau)\frac{d\:\mr{ln}\:P(k,\tau)}{d\:\mr{ln}\:k},
\eea
where $\alpha(\tau)$ and $\beta(\tau)$ are integrals involving background cosmological parameters and can be computed numerically \cite{2014Tides, 2016Tides}.
The above result only includes the leading order effect of the coupling between
the large-scale tidal field and small-scale density fluctuations. 
In reality the density field is quite nonlinear and involves all higher order 
interactions. 
The reconstructed density field would be biased when the theoretical description
of the nonlinear coupling in the above equation is not accurate.
This problem can be addressed using the transfer function calibrated from 
simulations \cite{2016Tides}.

The traceless tidal tensor $t_{ij}$ can be decomposed into five independently 
observable components ($\gamma_1$, $\gamma_2$, $\gamma_x$, $\gamma_y$, 
$\gamma_z$) \cite{2016Tides}.
We notice that the two transverse shear terms, 
\bea
\gamma_1=(\Phi_{L,11}-\Phi_{L,22})/2,\ \gamma_2=\Phi_{L,12},
\eea
which describe quadrupolar distortions in the tangential plane perpendicular 
to the line of sight, are less affected by peculiar velocities.
Thus, in the following computation we shall use them to perform reconstruction.
Once we have the tidal shear terms $\gamma_1$ and $\gamma_2$, the reconstructed
density field can be obtained by
\bea
\label{eq:rec}
\delta_r(\bmk)=\frac{Ak^2}{(k_1^2+k_2^2)^2}\bigg[(k_1^2-k_2^2)\gamma_1(\bmk)+2k_1k_2\gamma_2(\bmk)\bigg],
\eea
where $A$ is the normalization coefficient.
Since we only use two transverse tidal shear fields $\gamma_1$ and $\gamma_2$
for reconstruction, the change of the large-scale density field along the line 
of sight is inferred from the variations of $\gamma_1$ and $\gamma_2$ along 
the $z$-axis. 
The noise for the reconstructed density field is anisotropic in Fourier space.
The tidal reconstruction technique works best for modes in the high $\kpe$ 
and low $\kpa$ region, which can not be obtained from 21-cm surveys directly 
but contribute substantially to observables from other cosmological observations
as discussed above. Cosmic tidal reconstruction provides a new possible way 
to recover the lost radial modes and to improve the cross-correlation signals.

\subsection{Reconstruction algorithm}
In this subsection, we describe the tidal reconstruction method used in the 
next section.

\subsubsection{Reducing nonlinearities}

The first step is to smooth the nonlinear density field with a Gaussian kernel, 
\bea
\delta_R(\bmk)=W_R(\bmk)\delta(\bmk),
\eea
where 
\bea
W_R(\bm{k})=\mr{exp}(-k^2R^2/2),
\eea
which filters out small-scale structures.
The perturbative description of tidal coupling in Eq. (\ref{eq:pk}) is not valid
in the strong non-Gaussian regions. 
We need to smooth small-scale nonlinear structures to reduce nonlinearities.
Here, we take $R=1.25\ \mr{Mpc}/h$, which is close to the optimal filter scale
as demonstrated in Refs. \cite{2012Tides,2016Tides}.

The second step is to gaussianize the smoothed density field by taking a 
logarithmic transform or mapping the density fluctuations into a Gaussian 
distribution according to their density values. 
We shall use the latter method since after the simulated 21-cm foreground 
subtraction process, some of the density contrasts become smaller than $-1$,
which makes it hard to take the logarithmic transform, $\mathrm{ln}(1+\delta)$.

\subsubsection{Estimating tidal shear fields}

The coupling of the large-scale tidal field and small-scale density fluctuations
leads to a local anisotropy of quadratic statistics.
The tidal shear fields can be reconstructed by applying quadratic estimators to
the Gaussianized density field $\delta_g$ as
\begin{eqnarray}
&&\hat{\gamma}_1(\bmx)=\big[\delta_g^{w_1}(\bmx)\delta_g^{w_1}(\bmx)+
\delta_g^{w_2}(\bmx)\delta_g^{w_2}(\bmx)\big]/2, \nonumber \\
&&\hat{\gamma}_2(\bmx)=\big[\delta_g^{w_1}(\bmx)\delta_g^{w_2}(\bmx)\big],
\end{eqnarray}
where 
\bea    
\delta_g^{w_1}(\bmk)=i\hat{k}_1w(\bmk)\delta_g(\bmk),\ 
\delta_g^{w_2}(\bmk)=i\hat{k}_2w(\bmk)\delta_g(\bmk),
\eea
and the filter is
\bea
w(\bmk)=\frac{\sqrt{P(\bmk)f(\bmk)}}{P_\mr{tot}(\bmk)},
\eea
Here, $P_\mr{tot}(\bmk)$ is the total power spectrum of
the 21-cm density field which includes both the signal and noise.
The reconstruction of tidal shear fields is similar to the reconstruction of
lensing shear fields from 21-cm temperature fields.
The quadratic estimators presented above can be constructed using either 
the maximum likelihood method or the inverse variance weighting 
\cite{2008LensingEst,2010LensingEst,2012LensingEst}.

\subsubsection{Generating the density field}

After we get the tidal shear fields $\gamma_1$ and $\gamma_2$, the tidal 
reconstructed density field is given by Eq. (\ref{eq:rec}).
In general, the reconstructed field $\delta_r(\bmk)$ can be written as
\bea
\delta_r(\bmk)=C(\bmk)\delta(\bmk)+N(\bmk),
\eea
where $C(\bmk)=P_{\delta_r\delta}(\bmk)/P_{\delta}(\bmk)$, $\delta(\bmk)$ is 
the original  matter density field and $N(\bmk)$ includes the noises from
21-cm observation and from reconstruction. 
The factor $C(\bmk)$, often referred as the propagator, quantifies how much 
information of the original density distribution is reconstructed.
To get an unbiased measurement of the original density field, we can deconvolve 
the propagator $C(\bmk)$ from the reconstructed field,
\bea
\label{eq:delta}
\hat{\delta}_r=\delta_r(\bmk)/C(\bmk)=\delta(\bmk)+N(\bmk)/C(\bmk).
\eea
This factor can be computed by performing reconstruction with the simulated
observation mock data \cite{2017NR2,2017FM,2017NR3}.

\section{Implementation and results}
\label{sec:sim}

To test the performance of reconstruction, we run an ensemble of six $N$-body
simulations with the ${\tt CUBEP^3M}$ code \cite{2013CUBEP3M}. 
Each simulation involves $1024^3$ dark matter particles in a cubic box of side
length $1200\ \mr{Mpc}/h$. 
We use the snapshot at redshift $z=1$ and generate the dark matter density 
field on a $1024^3$ grid.
We could approximately use the dark matter density to represent the 21-cm 
source distribution, i.e., the neutral hydrogen.
This is a good approximation since the neutral hydrogen traces the total mass
distribution fairly well at low redshifts (see Refs. \cite{2017HI,2017Void}
for more discussions about the modeling of neutral hydrogen in the Universe).

There are several noises for 21-cm experiments we need to consider to model
the observed 21-cm signal from intensity mapping observations, including the 
astrophysical foreground, the receiver noise, and the shortest baseline for 
inteferometers.
